PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-13 paper/sec-pineappl.tex
Add small fixes
- hyphenation of 'Dittmaier'
- don't use small caps PineAPPL in titles, latex doesn't do that
- use pt of the sum of the two leptons instead of the Z boson
- use SIrange for percentage ranges

\subsection{The \textsc{PineAPPL} library}
\label{sec:library}

\textsc{PineAPPL} is a new library which stores phase-space weights of a Monte Carlo (MC) integration of a fixed-order calculation independently from the chosen PDFs.
The task of computing predictions for physical observables is therefore split
into two steps: 1) the generation of the \emph{grids}, that is the files in
which the MC weights are stored, and 2) the convolution of these grids with a set of PDFs.
The advantage of this method is that the time-consuming step 1) has to be performed only once, and that step 2) is reduced to a fast convolution of a given grid with one (or more) PDF set(s).

>>>>

\subsection{The PineAPPL library}
\label{sec:library}

\textsc{PineAPPL} is a new library which stores phase-space weights of a Monte Carlo (MC) integration of a fixed-order calculation independently from the chosen PDFs.
The task of computing predictions for physical observables is therefore split
into two steps: 1) the generation of the \emph{grids}, i.e.\ the files in
which the MC weights are stored, and 2) the convolution of these grids with a set of PDFs.
The advantage of this method is that the time-consuming step 1) has to be performed only once, and that step 2) is reduced to a fast convolution of a given grid with one (or more) PDF set(s).

### 2020-08-13 paper/sec-results.tex
General revision
Add small fixes
- hyphenation of 'Dittmaier'
- don't use small caps PineAPPL in titles, latex doesn't do that
- use pt of the sum of the two leptons instead of the Z boson
- use SIrange for percentage ranges

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Top-quark pair production.}
\paragraph{Experimental measurements and process features.}
We select the single-differential distribution in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml} (see
also refs.~\cite{Pagani:2016caq,Czakon:2017wor} for studies related to the photon density) and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. These orders make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contributions $\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$, which have been shown to be negligible~\cite{Czakon:2017wor,Frederix:2018nkq}.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
7 (37) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Z-boson production with non-zero transverse momentum.}

\paragraph{Experimental measurements and process features.}
We select the single-differential transverse momentum distribution of the
the Z boson, $p_\mathrm{T}^\mathrm{Z}$, measured by the CMS experiment at a
centre-of-mass energy of \SI{13}{\tera\electronvolt}~\cite{Sirunyan:2019bzr}.
This measurement, which has not been included in any PDF determination so far,
shows very low experimental uncertainties (at the percent level or below).
EW corrections are therefore expected to be essential to achieve a good
description of it, and to constrain accurately the PDFs, together with
NNLO QCD corrections, which are already well
known~\cite{Gehrmann-DeRidder:2017mvr,Bizon:2019zgf}
{\bf MZ there are at least 5-6 other papers both
  by the Zurich group and the Argonne one, but on Z+jet rather than
  Z-pt predictions. Should we cite them?}
Analogous measurements,
from the ATLAS~\cite{Aad:2015auj} and CMS~\cite{Khachatryan:2015oaa}
experiments at a centre-of-mass energy of \SI{8}{\tera\electronvolt},
were partly included (upon
selection of an appropriate kinematic cut that excluded bins with large EW
corrections) in a dedicated study~\cite{Boughezal:2017nla},
in the NNPDF3.1 PDF set~\cite{Ball:2017nwa} and in variants of
the CT18 PDF set~\cite{Hou:2019efy}. In the QCD computation, we consider a
single LO contribution $\mathcal{O}(\alphas\alpha^2)$ and a single NLO
contribution $\mathcal{O}(\alphas^2\alpha^2)$. In the NLO QCD+EW computation,
we supplement these with another LO and NLO, which are
$\mathcal{O}(\alpha^3)$ and $\mathcal{O}(\alphas\alpha^3)$;
contributions of the order $\mathcal{O}(\alpha^4)$ are not considered (see ref.~\cite{Denner:2019zfp}).
EW corrections for this process were computed in
refs.~\cite{Kuhn:2005az,Denner:2011vu,Hollik:2015pja,Kallweit:2015dum}.
The process receives contributions from 101 (166) parton luminosities,
see appendix~\ref{app:lumis}.

@@

As in the case of Drell-Yan, we use a fixed value for the renormalisation and
factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$, where
$M_\mathrm{Z}$ is the $\mathrm{Z}$-boson mass. Consistently with the
experimental analysis, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|M_{\ell\bar\ell} - M_\mathrm{Z}| < \SI{15}{\giga\electronvolt}$,
$|y_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<p_\mathrm{T}^{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the
transverse momentum and pseudorapidity of each lepton, and for the invariant
mass, pseudorapidity and transverse momentum of the lepton pair.
We finally discard all bins with
$p_\mathrm{T}^\mathrm{Z}<\SI{20}{\giga\electronvolt}$ to avoid a kinematic
ergion where resummation effects are sizeable.

\paragraph{Numerical results.} In figure~\ref{fig:cmsZ13TeV} we report the
distribution differential in the transverse momentum of the $\mathrm{Z}$
boson. Also in this case, the \textsc{PineAPPL} result is well validated, as it
differs from the MC result by $\SI{0.5}{\permille}$ at most. The accuracy of the
theory or the choice of scale do not alter this conclusion.

@@

The process receives its leading contribution from $\mathrm{qg}$- and
$\bar{\mathrm{q}}\mathrm{g}$-initiated channels, which account for about 65\%
of the cross section, with some variations in the relative contributions from
individual quark and atiquark flavours as the transverse momentum of the
$\mathrm{Z}$ boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
2\% -- 7\% as the transverse momentum of the $\mathrm{Z}$ boson increases.
The size of this shift is usually as large as the data uncertainty, and
exceeds it at large values of $p_\mathrm{T}^\mathrm{Z}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit to this data set.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Top-quark pair production.}
\label{sec:toppair}

\paragraph{Experimental measurements and process features.}
We select the single-differential distribution in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml} (see
also refs.~\cite{Pagani:2016caq,Czakon:2017wor} for studies related to the photon density) and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. These orders make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contributions $\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$, which have been shown to be negligible~\cite{Czakon:2017wor,Frederix:2018nkq}.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
7 (37) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Z-boson production with non-zero transverse momentum.}
\label{sec:Zpt}

\paragraph{Experimental measurements and process features.}
We select the single-differential transverse momentum distribution of the
sum of the two leptons (the \enquote{Z boson}), $p_\mathrm{T}^{\ell \bar{\ell}}$, measured by the CMS experiment at a
centre-of-mass energy of \SI{13}{\tera\electronvolt}~\cite{Sirunyan:2019bzr}.
This measurement, which has not been included in any PDF determination so far,
shows very low experimental uncertainties (at the percent level or below).
EW corrections are therefore expected to be essential to achieve a good
description of it, and to constrain accurately the PDFs, together with
NNLO QCD corrections, which are already well
known~\cite{Gehrmann-DeRidder:2017mvr,Bizon:2019zgf}
{\bf MZ there are at least 5-6 other papers both
  by the Zurich group and the Argonne one, but on Z+jet rather than
  Z-pt predictions. Should we cite them?}
Analogous measurements,
from the ATLAS~\cite{Aad:2015auj} and CMS~\cite{Khachatryan:2015oaa}
experiments at a centre-of-mass energy of \SI{8}{\tera\electronvolt},
were partly included (upon
selection of an appropriate kinematic cut that excluded bins with large EW
corrections) in a dedicated study~\cite{Boughezal:2017nla},
in the NNPDF3.1 PDF set~\cite{Ball:2017nwa} and in variants of
the CT18 PDF set~\cite{Hou:2019efy}. In the QCD computation, we consider a
single LO contribution $\mathcal{O}(\alphas\alpha^2)$ and a single NLO
contribution $\mathcal{O}(\alphas^2\alpha^2)$. In the NLO QCD+EW computation,
we supplement these with another LO and NLO, which are
$\mathcal{O}(\alpha^3)$ and $\mathcal{O}(\alphas\alpha^3)$;
contributions of the order $\mathcal{O}(\alpha^4)$ are not considered (see ref.~\cite{Denner:2019zfp}).
EW corrections for this process were computed in
refs.~\cite{Kuhn:2005az,Denner:2011vu,Hollik:2015pja,Kallweit:2015dum}.
The process receives contributions from 101 (166) parton luminosities,
see appendix~\ref{app:lumis}.

@@

As in the case of Drell-Yan, we use a fixed value for the renormalisation and
factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$, where
$M_\mathrm{Z}$ is the Z-boson mass. Consistently with the
experimental analysis, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|M_{\ell\bar\ell} - M_\mathrm{Z}| < \SI{15}{\giga\electronvolt}$,
$|y_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<p_\mathrm{T}^{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the
transverse momentum and pseudorapidity of each lepton, and for the invariant
mass, pseudorapidity and transverse momentum of the lepton pair.
We finally discard all bins with
$p_\mathrm{T}^{\ell \bar{\ell}}<\SI{20}{\giga\electronvolt}$ to avoid a kinematic
region where resummation effects are sizeable.

\paragraph{Numerical results.} In figure~\ref{fig:cmsZ13TeV} we report the
distribution differential in the transverse momentum of the Z
boson. Also in this case, the \textsc{PineAPPL} result is well validated, as it
differs from the MC result by $\SI{0.5}{\permille}$ at most. The accuracy of the
theory or the choice of scale do not alter this conclusion.

@@

The process receives its leading contribution from $\mathrm{qg}$- and
$\bar{\mathrm{q}}\mathrm{g}$-initiated channels, which account for about \SI{65}{\percent}
of the cross section, with some variations in the relative contributions from
individual quark and antiquark flavours as the transverse momentum of the
Z boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
\SIrange{2}{7}{\percent} as the transverse momentum of the Z boson increases.
The size of this shift is usually as large as the data uncertainty, and
exceeds it at large values of $p_\mathrm{T}^{\ell \bar{\ell}}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit to this data set.
